Containerization of CMS Applications with Docker

### 2015-04-08 paper.tex
first commit

Docker is an open platform to build, ship and run distributed applications. It is a container-based virtualisation framework that uses Linux containers (LXC) as the core technology, with a large set of tools allowing for large scale container handling, shipping and deployment. It allows for easy container contextualization, and for inter-container communications. Docker utilization is becoming the de-facto standard for container based virtualisation, and is the recommended solution endorsed by most heavy weight IT providers. Some examples are Spotify, that uses Docker for continuous delivery, service testing and deployment; Baidu, that uses Docker as a Platform-as-a-Service (PaaS), profiting of its flexibility for many framework and applications; eBay, that uses Docker for its easy application deployment, and for continuous integration process; and many more. Recently, Docker has reached 120k lines of code, about 10k commits and roughy 600 core contributors, of which about 95\% work outside Docker.

\section{Docker as a tool for light virtualization}

Virtualization can be achieved in several ways, depending on the desired (or acceptable) level of invasiveness. Setting aside solutions when the whole system hardware is emulated (like QEMU), the industry reference technologies are full system virtualization (OpernStack, OpenNebula), where each virtualised machine has its own running kernel, and kernel based process isolation, which has been a possibility in Linux since the beginning, via \emph{chroot}, and more recently via \emph{c-groups} and projects like the Linux Container (\emph{LXC}). The former technology has a broader scope, being able to utilize different operating systems on the same physical host (e.g. mixing Windows and Linux machines), but pays a price in terms of resource utilization (mainly RAM, to a smaller extent CPU reduced efficiency) and access to high performance devices; it is considered an overkill in our field, where more or less all the scirntific computation are carried out on Linux machines. Linux Containers, at the heart of the Docker technology, instead of simulating complete machines, just provede process separation and sand-boxing over a common kernel all the machines share. This is particularly adequate in an environment where system standardization has already been achieved to a large extent, for example by the adoption of WLCG-provided Middleware.
Docker provides CPU, memory and filesystem isolation, and can run different processes on the same kernel but with completely different runtime (e.g. one can run SLC5 on SLC6 or on Ubuntu).

Docker adds on top of LXC a complete series of tools for image storing, deployment, testing and contextualization, as well as for container handling (start/stop/...) and communications; derived images can be based on exiting images, with Docker taking care to save only the deltas between them. It also add a public repository (Docker Hub, =hub.docker.com=), which allows for safe image storing and easy deployment for open software projects, which provides already cooked images for most of the widespread linux based operating systems.

@@

\section{``Dockerizing'' a computing site}
Virtualization technologies are the best choice when dealing when dealing with users with diverse needs in term of computing resources. While Linux is the de-facto standard for most scientific uses, precise hardware and software needs can still vary to the extent of not allowing for a catch-all solution. Software from previous generation experiment is often not validated for the latest-greatest Linux versions, for example, and thus cannot coexist with recent experiments which instead prefer the best performance coming with newer releases.

@@

\begin{itemize}
\item GPFS is used to serve data for all the use-cases supported on-site;
\item AFS is used for user areas; 
\item LSF is used for resource access, also interactive.
\end{itemize}

The main issue in Pisa recently was to understand how to provision the correct environment to all these diverse resources. Virtual Machines (e.g. OpenStack) were considered as an option, but
Pisa also has older machines which are low in RAM. Infiniband connectivity, moreover, seems to loose performance in a completely virtualised environment. The solution identified up to recently so far was a very light virtualisation via chroot. Every host machine (with the very latest OpenSUSE kernel) was starting sand-boxed machines as tar files containind complete SL6 (or SLx) systems,  via chroot. Site-wide file-systems (CVMFS/GPFS/AFS) were mounted by the host, and then seen by the machines as local file systems using the emph{bind}  tool.
Every machine had pre-installed as many tar files as the possible environments are: these could be un-tarred and ``started'' on demand via a set of scripts, basically forcing the LSF pool the machine lives in. This solution works and it has been in production for Pisa since 3 years.

On the other hand, Docker is to be a viable solution, too. It has the same overhead (virtually 0) as chroot, and comes with the added plus of easier image storing, deployment and control. Basically, three are the main points in favour of Docker adoption in Pisa: 
\begin{itemize}
\item no more tar files and a complicate management;
\item the adoption of a git-like image management, where every action on the repository is properly logged, and you can bransh, pull, etc;
\item the very easy deployment of a local image repository, to avoid exposure of sensitive information;
\item the extremely easy conversion of an existing chroot's tar image to a Docker image, via a single command:
\begin{verbatim}
\end{verbatim}
    sudo tar -C ExistingChrootBaseDir -c . | sudo docker import - PisaWN
\end{itemize}

A Tier-2 on-demand approach was initially planned, where no dependency on the underlying physical architecture was foreseen, but currently it has been decided to privilege ese of deplyment and performance over

attractive, but site performances for a site in production are most important at the moment. For example, Pisa still decided to mount CVMFS/GPFS/AFS on the host and pass this via ``-v''. This is not something you can ask e.g. in opportunistic sites, but this {\em i)} avoids running docker privileged, and {\em ii)}
avoids multiple copies of caches (CVMFS/AFS) if more than one container is running. LSF is also in the container, so it automatically connects the machine to the proper LSF queue, with a start command as follows:
%
\begin{verbatim}
    docker run -v /cvmfs:/cvmfs -v /afs:/afs -v /gpfs/ddn:/gpfs/ddn
    -v /chrootlfs/home:/home/grid -d
    -t localregistry.pi.infn.it:5000/enricomazzoni/testwn:0.3
    /etc/sysconfig/docker-pi/start
\end{verbatim}
%
Pisa moved ~10\% of the WNs to Docker so far, all work done in 7-10 days. No user actually realised the difference. The rest of the centre will migrate as well in the next few weeks. At this point, the idea would be to move all the services used by CMS via Docker, e.g.: {\em i)} Squid: just a Linux standard machine; {\em ii)} CEs: Pisa moved one and it’s working; {\em iii)} UIs: easy, similar to WNs; {\em iv)} PhEDEx (transfer agents on a UI), it should be as easy as a UI; {\em v)} an xrootd redirector (data federations): Pisa moved it from chroot and it is already running now with Docker.

@@

In summary, Docker offers interesting opportunities to CMS. Itcan ship entire CMSSW distributions and do benchmarking easily. It offers plenty of desired features and flexibility at a site level, and all this at no cost in terms of performances (as from the tests done so farm, at least).

@@

%%%%%%
%\begin{figure}
%\begin{center}
%\includegraphics[width=.8\textwidth]{plots/datasources.pdf}
%\caption{bla bla}
%\label{FIG:datasources}
%\end{center}
%\end{figure}
%%%%%%
%

\begin{thebibliography}{99}
\bibitem{DockerGiulio} CMS examples with Docker: \texttt{http://github.com/cms-sw/cms-docker}
\bibitem{parfullcms} ParFullCMS, \texttt{https://github.com/cms-sw/cms-docker/tree/master/parfullcms}

>>>>

Docker~\cite{docker} is an open platform to build, ship and run distributed applications. It is a container-based virtualisation framework that uses Linux containers (LXC) as the core technology, with a large set of tools allowing for large scale container handling, shipping and deployment. It allows for easy container contextualization, and for inter-container communications. Docker utilisation is becoming the de-facto standard for container based virtualisation, and is the recommended solution endorsed by most heavy weight IT providers. Some examples are Spotify, that uses Docker for continuous delivery, service testing and deployment; Baidu, that uses Docker as a Platform-as-a-Service (PaaS), profiting of its flexibility for many framework and applications; eBay, that uses Docker for its easy application deployment, and for continuous integration process; and many more. Recently, Docker has reached 120k lines of code, about 10k commits and roughy 600 core contributors, of which about 95\% work outside Docker.

\section{Docker as a tool for light virtualisation}

Virtualisation can be achieved in several ways, depending on the desired (or acceptable) level of invasiveness. Setting aside solutions when the whole system hardware is emulated (like QEMU), the industry reference technologies are full system virtualisation (OpenStack~\cite{openstack}, OpenNebula~\cite{opennebula}), where each virtualised machine has its own running kernel, and kernel based process isolation, which has been a possibility in Linux since the beginning, via \emph{chroot}, and more recently via \emph{c-groups} and projects like the Linux Container (\emph{LXC}~\cite{lxc}). The former technology has a broader scope, being able to utilize different operating systems on the same physical host (e.g. mixing Windows and Linux machines), but pays a price in terms of resource utilisation (mainly RAM, to a smaller extent CPU reduced efficiency) and access to high performance devices; it is considered an overkill in our field, where more or less all the scientific computation are carried out on Linux machines. Linux Containers, at the heart of the Docker technology, instead of simulating complete machines, just provede process separation and sand-boxing over a common kernel all the machines share. This is particularly adequate in an environment where system standardisation has already been achieved to a large extent, for example by the adoption of WLCG-provided Middleware.
Docker provides CPU, memory and filesystem isolation, and can run different processes on the same kernel but with completely different runtime (e.g. one can run SLC5 on SLC6 or on Ubuntu).

Docker adds on top of LXC a complete series of tools for image storing, deployment, testing and contextualization, as well as for container handling (start/stop/...) and communications; derived images can be based on exiting images, with Docker taking care to save only the deltas between them. It also add a public repository (Docker Hub, {\verb hub.docker.com}), which allows for safe image storing and easy deployment for open software projects, which provides already cooked images for most of the widespread linux based operating systems.

@@

\section{``Dockerizing'' a computing site}
Virtualisation technologies are the best choice when dealing when dealing with users with diverse needs in term of computing resources. While Linux is the de-facto standard for most scientific uses, precise hardware and software needs can still vary to the extent of not allowing for a catch-all solution. Software from previous generation experiment is often not validated for the latest-greatest Linux versions, for example, and thus cannot coexist with recent experiments which instead prefer the best performance coming with newer releases.

@@

\begin{itemize}
\item GPFS~\cite{gpfs} is used to serve data for all the use-cases supported on-site;
\item AFS~\cite{afs} is used for user areas; 
\item LSF~\cite{lsf} is used for resource access, also interactive.
\end{itemize}

The main issue in Pisa recently was to understand how to provision the correct environment to all these diverse resources. Virtual Machines (e.g. OpenStack) were considered as an option, but
Pisa also has older machines which are low in RAM. Infiniband connectivity, moreover, seems to loose performance in a completely virtualised environment. The solution identified up to recently so far was a very light virtualisation via chroot. Every host machine (with the very latest OpenSUSE kernel) was starting sand-boxed machines as tar files containing complete SL6 (or SLx) systems,  via chroot. Site-wide file-systems (CVMFS~/cite{cvmfs}/GPFS/AFS~\cite{afs}) were mounted by the host, and then seen by the machines as local file systems using the emph{bind}  tool.
Every machine had pre-installed as many tar files as the possible environments are: these could be un-tarred and ``started'' on demand via a set of scripts, basically forcing the LSF pool the machine lives in. This solution works and it has been in production for Pisa since 3 years.

On the other hand, Docker is to be a viable solution, too. It has the same overhead (virtually 0) as chroot, and comes with the added plus of easier image storing, deployment and control. Basically, three are the main points in favour of Docker adoption in Pisa: 
\begin{itemize}
\item no more tar files and a complicate management;
\item the adoption of a git-like image management, where every action on the repository is properly logged, and you can branch, pull, etc;
\item the very easy deployment of a local image repository, to avoid exposure of sensitive information;
\item the extremely easy conversion of an existing chroot's tar image to a Docker image, via a single command:
\begin{verbatim}
sudo tar -C ExistingChrootBaseDir -c . \
| sudo docker import - PisaWN
\end{verbatim}
\end{itemize}

A Tier-2 on-demand approach was initially planned, where no dependency on the underlying physical architecture was foreseen, but currently it has been decided to privilege ese of deployment and performance over flexibility. Pisa decided hence to mount CVMFS/GPFS/AFS on the host and pass them via ``-v'', instead of relying on the image mounting them via userland tools like Parrot;
The solution also allows for the sharing of CVMFS/AFS caches in case of multiple machines running, and does not force the containers to run in privileged mode, which would not be possible in (e.g.) opportunistic sites.
LSF  runs within the container, and automatically connects the machine to the proper LSF queue, with a start command as follows. all in all, a Worker Node instance is started just via
%
\begin{verbatim}
docker run -v /cvmfs:/cvmfs -v /afs:/afs \
-v /gpfs/ddn:/gpfs/ddn \
-v /chrootlfs/home:/home/grid -d \
-t localregistry.pi.infn.it:5000/enricomazzoni/testwn:0.3 \
/etc/sysconfig/docker-pi/start
\end{verbatim}

The complete workflow of operations is shown if Fig.\ref{fig:dockerPisa}: admin's action selects and starts the appropriate container, which then joins the correct LSF queue and becomes available for processing (via batch system and interactive).

%
Pisa moved ~10\% of the WNs to Docker so far, following an R\&D process of just a couple of weeks. No user actually realised the difference. The rest of the centre will migrate as well in the next few weeks. At this point, the idea would be to move all the services used by CMS via Docker:
\begin{itemize}
\item Squid: just a Linux standard machine;
\item Computing Elements (CEs): one already moved without problems, starting from a chtooed test we had;
\item User Interfacess (UIs): similar to WNs, can share most of their image; 
\item PhEDEx (CMS transfer agents on a User Interfac): a simple modification of the latter;
\item Xrootd redirectors (for CMS data federations): Pisa moved it from chroot and it is already running now with Docker.
\end{itemize}

@@

In summary, Docker offers interesting opportunities to CMS. It can ship entire CMSSW distributions and do benchmarking easily. It offers plenty of desired features and flexibility at a site level, and all this at no cost in terms of performances (as from the tests done so farm, at least).

@@

\begin{figure}
\begin{center}
\includegraphics[width=.8\textwidth]{dockerpisa.pdf}
\caption{Workflow from admin's action to a machine accessible via LSF.}
\label{fig:dockerPisa}
\end{center}
\end{figure}
%

\begin{thebibliography}{99}
\bibitem{docker} Docker Web Site, \texttt{http://www.docker.com/}.
\bibitem{openstack}OpenStack, \texttt{https://www.openstack.org/}.
\bibitem{opennebula}OpenNebula, \texttt{http://opennebula.org/}.
\bibitem{lxc}Linux Containers, \texttt{http://en.wikipedia.org/wiki/LXC}.
.\bibitem{gpfs} GPFS: A Shared-Disk File System for Large Computing Clusters, Proceedings of the FAST 2002 Conference on File and Storage Technologies Monterey, California, USA January 28-30, 2002.
\bibitem{afs} OpenAFS, \texttt{http://www.openafs.org/}.
\bibitem{cvmfs} CVMFS, \texttt{http://cernvm.cern.ch/portal/filesystem}.
\bibitem{lsf} LSF, \texttt{http://www-03.ibm.com/systems/platformcomputing/products/lsf}.
\bibitem{parrot}The Parrot Virtual File System, \texttt{The Parrot Virtual File System}.
\bibitem{DockerGiulio} CMS examples with Docker: \texttt{http://github.com/cms-sw/cms-docker}.
\bibitem{parfullcms} ParFullCMS, \texttt{https://github.com/cms-sw/cms-docker/tree/master/parfullcms}.
